Quantifying the importance of a tree prior in Bayesian phylogenetics

### 2019-04-15 article.tex
Updates from Overleaf

The second step (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 3a)
selects inference models from a set of experiments. 
We define an experiment $X$ as the combination of an inference model $I$
and the conditions $C$ to actually use it in the inference step.
There are multiple possible setups for $C$. 
If the attached inference model $I$ is the generative model, then it should always be selected. If, instead, the attached inference model is a candidate model the user can either let \verb;pirouette; select it as the best among all the candidate models, or force its selection (regardless
of it being the best).
When $I$ is set to be the generative model,
it will use the 
same site and clock model used in the alignment simulation,
and the same tree prior used in the creation of the given phylogeny. 
A 'best' model is defined as the inference model with
the highest evidence (a.k.a. marginal likelihood) for that inference model, 
given the alignment simulated in the previous step.
The evidence of an inference model is estimated by nested 
sampling [\cite{maturana2018model}], using the \verb;NS; BEAST2 package.
Scripted use of BEAST2 packages is only possible under under Linux and Mac.
Windows systems can do the model comparison for shorter DNA sequences
using the web interface of \verb;mcbette; [\cite{mcbette}].

@@

The fourth step quantifies the inference error made, by removing
the burn-in, then applying an error statistic.
The burn-in is the first phase of the MCMC,
where it samples an unrepresentative part of state space,
and is commonly discarded. By default, \verb;pirouette; 
uses removes the first 10\% of the posterior.
From the remaining posterior, \verb;pirouette; 
creates an error distribution, by measuring the difference
between the true tree and each of the posterior 
trees (Fig.~\ref{fig:pipeline}, 4a $\rightarrow$ 5a).
The default way to quantify the difference between two phylogenies
is the nLTT statistic (\cite{janzen2015approximate}), but any 
user-defined error statistic can be used.

@@

An optional step is to generate a 'twin tree' $\tau_{I}$
(Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 1b),
that will be analyzed in the same way as the true tree.
Any phylogeny $\tau$ is the combination of 
branching times $\Vec{t}$ and topology $\psi$.
Starting from a non-standard generative prior $t_p = \mathit{P}$,
we can observe a given tree $\tau_{\mathit{P}}$,
with branching times $\Vec{t}_{\mathit{P}}$ and 
topology $\psi_{\mathit{P}}$, 
but we cannot use the tree prior in Bayesian inference.
The twinning process, T, in accordance with a chosen inference 
model $t_p \in \mathit{I}$, creates a tree $\tau_{\mathit{I}}$
with branching times $\Vec{t}_{\mathit{I}}$ while preserving 
topology $\psi_{\mathit{P}}$:

>>>>

The second step (Fig.~\ref{fig:pipeline}, 2a $\rightarrow$ 3a)
selects inference models from a set of experiments. 
We define an experiment $X$ as the combination of an inference model $I$
and the conditions $C$ to actually use it in the inference step.
There are multiple possible setups for $C$. 
If the attached inference model $I$ is the generative model, then it should always be selected. If, instead, the attached inference model is a candidate model the user can either let \verb;pirouette; select it as the best among all the candidate models, or force its selection (regardless
of it being the best).
When $I$ is set to be the generative model,
it will use the 
same site and clock model used in the alignment simulation
and the same tree prior used in the creation of the given phylogeny. 
A 'best' model is defined as the inference model with
the highest evidence (a.k.a. marginal likelihood) for that inference model, 
given the alignment simulated in the previous step.
The evidence of an inference model is estimated by nested 
sampling [\cite{maturana2018model}], using the \verb;NS; BEAST2 package.
Scripted use of BEAST2 packages is only possible under under Linux and Mac.
Windows systems can do the model comparison for shorter DNA sequences
using the web interface of \verb;mcbette; [\cite{mcbette}].

@@

The fourth step quantifies the inference error made, by removing
the burn-in, then applying an error statistic.
The burn-in is the first phase of the Markov chain Monte Carlo (MCMC) run,
in which it samples an unrepresentative part of state space,
and is commonly discarded. By default, \verb;pirouette; 
removes the first 10\% of the posterior.
From the remaining posterior \verb;pirouette; 
creates an error distribution, by measuring the difference
between the true tree and each of the posterior 
trees (Fig.~\ref{fig:pipeline}, 4a $\rightarrow$ 5a).
The default way to quantify the difference between two phylogenies
is the nLTT statistic (\cite{janzen2015approximate}), but any 
user-defined error statistic can be used.

@@

An optional step is to generate a 'twin tree' $\tau_{I}$
(Fig.~\ref{fig:pipeline}, 1a $\rightarrow$ 1b),
that will be analyzed in the same way as the true tree.
Any phylogeny $\tau$ is the combination of 
branching times $\Vec{t}$ and topology $\psi$.
Starting from a non-standard generative prior $t_p = \mathit{P}$,
we can observe a given tree $\tau_{\mathit{P}}$,
with branching times $\Vec{t}_{\mathit{P}}$ and 
topology $\psi_{\mathit{P}}$, 
but we cannot use the tree prior in Bayesian inference.
The twinning process T, in accordance with a chosen inference 
model $t_p \in \mathit{I}$, creates a tree $\tau_{\mathit{I}}$
with branching times $\Vec{t}_{\mathit{I}}$ while preserving the 
topology $\psi_{\mathit{P}}$:
